### 2009-11-27 docs/marcello-dissertation/thesis/chapter10-appendix.tex
Major revision of the thesis structure, and review of chapters, moving chapter 7 to the appendix, and adding the first revision of the abstract in the preamble.
Chap 3. Reviewed the structure, misspelled words, sentences, etc;

Chap 4. Changes to the structure of the case study. The documentation regarding netbeams was pushed to the last chapter: appendix;

Chap 5. The selection of the technology follows the review with Prof. Puder. One section describes each of the technologies briefly, and details their characteristics during the analysis of each of the taxonomies;

Chap 6. Describes the imp

\section{Collected Sensor Data as the DSP Message Content}

@@

@@

@@

\lstset{label=file:mongodb-ysi-data-format,caption=JSON representation of the
data produced by a YSI sensor on mongoDB}
\begin{lstlisting}
            {
             "_id" :  ObjectId( "d36f4007b7e7ac4a03c60000")  , 
             "sensor_ip_address" : "192.168.0.136", 
             "message_id" : "7b6624d6-0ca1-4cba-a343-f166e88da73b",
             "transaction_time" : 1252845473412 , 
             "fact_time" : 1252845346000,
             "latitude" : 12.3450, 
             "longitude" : -123.4456 , 
             "data" : {
                       "temperature" : 45.01, 
                       "sp_condition" : 37.6, 
                       "condition" : 145.8, 
                       "resistence" : 159.77, 
                       "salinitude" : 0.0, 
                       "pressure" : 0.391, 
                       "depth" : 0.46, 
                       "ph" : 5.64, 
                       "pH_mv" : -62.1, 
                       "odo_sat" : 89.7, 
                       "odo_condition" : 59.34, 
                       "turbidity" : 0.0,
                       "battery" : 9.4
                      }
            }
\end{lstlisting}

>>>>

\section{Collected Sensor Data as the DSP Message Content}
\label{sec:dsp-payload-implementation}

@@

\chapter{DSP Data Persistence: Component Implementation}
\label{sec:dsp-data-persistence-implementation}

This section shows the implementation of the DSP Data Persistence, which
follows the specifications of the DSP Components in the previous chapter,
along with the setup of the mongoDB, the chosen database system that models
data in a type of key-value pair model. Moreover, this chapter also describes
the experiments conducted for the persistence.

The implementation of the DSP Data Persistence component, as well as the
implementation of the experiments, were developed and managed using the DSP a
reserved development branch at the Subversion \cite{subversion} repository
located at the Google Code website. In this way, anyone can checkout the entire
source-code from the branch
\textit{https://netbeams.googlecode.com/svn/branches/marcello/persistence},
as shown in Figure \ref{fig:dsp-data-persistence-dir}. In this way, the
recommended checkout branch directory will give the remaining directory
structure as summarized:

\begin{itemize}
  \item \textbf{NEBEAMS-DIR/}: the directories under the subdirectory
  ``persistence'';
  \item \textbf{DEV-VERSION}: the directory of the current development
  version for the DSP; That is, ``\textbf{NEBEAMS-DIR}/versions/v2/''
  \item \textbf{PERSISTENCE-DIR}: the directory of the DSP component, that is,
  ``\textbf{DEV-VERSION}/apps/osgi-intro-bundles/dsp/DSPDataPersistence''.
\end{itemize}

\begin{figure}[!b]
  \centering
  \includegraphics[scale=0.6]{../diagrams/dsp-data-persistence-dir}
  \caption{The DSP Data Persistence Directory Structure}
  \label{fig:dsp-data-persistence-dir}
\end{figure}

The DSP Platform defines its own run-time directory, which organizes the
artifacts used by the osgi-intro Framework, as well as the its own configuration
descriptor and matcher, as described in the previous section.

\begin{itemize}
  \item \textbf{RUNTIME-DIR}: the main directory for the execution of the DSP
Platform, with all the necessary osgi-intro-related artifacts are placed, as shown in
Figure \ref{fig:dsp-runtime-dir}.
\end{itemize}

\begin{figure}[!t]
  \centering
  \includegraphics[scale=0.5]{../diagrams/dsp-runtime-dir}
  \caption{The DSP Data Persistence Directory Structure}
  \label{fig:dsp-runtime-dir}
\end{figure}

\section{DSP Platform Deployment}

As described in the previous chapter, the DSP Data Persistence Component was
implemented using the Java Programming language, on top of the osgi-intro Framework
API. This section describes the setup process of the implementation, as well as
the artifacts use during the implementation, which will be listed in the
appendix section.

The structure of the artifacts implemented for the DSP Data Persistence
Component follow the conventions of the NetBEAMS implementation, where the main
directory \textbf{PERSISTENCE-DIR}, depicted by Figure
\ref{fig:dsp-data-persistence-dir-checkedout}, holds each of the development
artifacts.

\begin{itemize}
  \item \textbf{DSPDataPersistence}: Main directory with the build.xml (Listing
  \ref{file:dsp-build.xml}) and other Eclipse-related artifacts. The other
  major directories are also listed:
  \item \textbf{META-INF}: this directory contains the descriptor file for
  the osgi-intro Framework, this this case the MANIFEST.MF;
  \item \textbf{src}: the main directory structure for the source-code
  implemented. Note that it follows the Java specification for packaging, and
  therefore, lists the package org.netbeams.dsp.persistence as the main
  package, including the layers controller, model and osgi-intro.
\end{itemize}

\begin{figure}[!h]
  \centering
  \includegraphics[scale=0.5]{../diagrams/dsp-data-persistence-dir-checkedout}
  \caption{The DSP Data Persistence Directory Structure}
  \label{fig:dsp-data-persistence-dir-checkedout}
\end{figure}

\subsection{osgi-intro Deployment Process}

Since each NetBEAMS component is managed by an osgi-intro component and its
infrastructure, this section describes the basic functionality of the osgi-intro
platform \cite{osgi-intro}, a framework was conceived to support modularity in
terms resources-limited environments such mobile devices and vehicles, but it
was first widely deployed on Eclipe IDE\footnote{Integrated Development
Environment}, because it promotes the use of reusable loosely-coupled modules
using the Java Programming Language \cite{java}. The most basic layer of the
osgi-intro Framework depicted in Figure \ref{fig:layering-osgi-intro} are as follows:

\begin{itemize}
  \item \textbf{Module Layer}: manages the osgi-intro bundles deployed on the osgi-intro
  Platform, providing the necessary "wiring" of the components. In other words,
  the modules, called osgi-intro Bundles, can export and/or import packages in the
  level of a Java Class managed by the osgi-intro Framework;
  \item \textbf{Service Layer}: resposible for the interoperability between 2
  or more bundles, enabling the bundles to register services offered by
  the its specification;
  \item \textbf{Execution Layer}: executes the bundles, managing and changing
  their life-cycle through the bundle execution.
\end{itemize}

\begin{figure}[!h]
  \centering
  \includegraphics[scale=0.5]{../diagrams/layering-osgi}
  \caption{The osgi-intro Framework Layers}
  \label{fig:layering-osgi-intro}
\end{figure}

NetBEAMS uses the concept of the modularization through the use of the the
Producer-Consumer paradigm described in the previous section. Since the DSP
Components are essentially osgi-intro bundles, the interoperability between them are
given in the implementation and reuse existing bundles in the osgi-intro Framework
and the DSP Platform, described by an artifact called MANIFEST.MF as seen in
listing \cite{file:osgi-intro-manifest}. In general, an osgi-intro bundle must
provide specifications that describes the module to be published into the osgi-intro
Framework. In this way, the main properties of the osgi-intro MANIFEST.MF artifact can
used by the DSP Data Persistence can be summarized as follows:

\begin{itemize}
  \item \textbf{Bundle-Activator}: The name of the instance of an osgi-intro Activator
  class, responsible to manage the bundle. The implemented class
  ``DataPersistenceActivator'' provides the activation mechanisms for the
  services provided by the DSP Data Persistence Component;
  \item \textbf{Bundle-ClassPath}: the necessary Java Jars list needed to run
  the bundle. It lists the mongoDB Java driver as one of the required
  dependencies deployed in the same package;
  \item \textbf{Import-Package}: the Java Packages needed by the DSP Data
  Persistence Component Bundle. These Java Packages must be provided by other
  DSP components, forcing the execution to be dependent on those package.
  They are provided through a ``Export-Package'' section in other osgi-intro
  bundles and, as it is shown in listing \ref{file:osgi-intro-manifest}, they
  are from different packages.
\end{itemize}

Once the osgi-intro bundle is installed into the osgi-intro Platform, it will be managed by
the osgi-intro Execution layer and change the bundle state according to its
lify-cycle, as described in the previous chapter. Therefore, the packaging of
the osgi-intro source-code is done using a JAR\footnote{Java Archival Repository}
artifact specification \cite{java-tutorial}. In this way, the DSP Data
Platform osgi-intro bundle JAR is created by the Apache ANT build script
\cite{apache-ant} in Listing \ref{file:dsp-build.xml}. The task
``dsp-data-persistence.all'' compiles all the source-code created from the
design of the previous chapter and packages everything into the artifact
``DSPDataPersistence-x.x.jar'', where x.x is the version of the bundle. Figure
\ref{fig:dsp-runtime-dir} shows the DSP Data Persistence file under the
dirctory ``\textbf{RUNTIME-DIR}/deployment/''.

\subsection{Adding DSP Data Persistence into DSP Platform }

Once the new component was developed, the DSP Data Persistence component was
deployed by editing two main descriptors, as well as the optional bootstrap
message, as described in chapter 5:

\begin{itemize}
  \item \textbf{config.xml}: the DSP Data Persistence is added to the DSP
  Platform adding an entry for the component;
  \item \textbf{matcher\underline{ }config.xml}: the addition of the rules that
  filters the messages to the DSP Data Persistence;
  \item \textbf{DSPDataPersistence\underline{ }001.xml}: depicts the bootstrap
  message for the DSP Data Persistence Component, as shown in Listing 
  \ref{file:dsp-data-pers-bootstrap.xml}.
\end{itemize}

The first configuration artifact enables NetBEAMS to initialize contain the
descriptor of the component, as shown in Listing \ref{file:dsp-config.xml}.
The component is added with the highest priority, since it contains
dependencies to other DSP Components. 
\subsection{Starting the DSP Platform}

Once the DSP Platform automatically installs all the components described by
the configuration descriptor, the container verifies the dependencies and
starts each of the DSP components of Figure \ref{fig:knopflerfish-execution},
the osgi-intro Framework container Knopflerfish \cite{knopflerfish}.

\begin{figure}[!t]
  \centering
  \includegraphics[scale=0.65]{../diagrams/knopflerfish-execution}
  \caption{The execution of the Knopflerfish Container}
  \label{fig:knopflerfish-execution}
\end{figure}

\subsection{Execution Logs}

\section{mongoDB Deployment}

Given that Document-Oriented Model makes a good candidate persist sensors'
properties and the recently enumerated list technologies in the previous
article DSPDataPersistence, the open-source project called mongoDB was chosen
for the evaluation on our case study, the netBEAMS DSP Platform.

mongoDB supports storage based on collections of data, stored using BSON, a
binary representation the JSON data representation format, including dynamic
queries and indexing support. As it's stated in their web site, mongoDB
"bridges the gap between key/value stores (which are fast and highly scalable)
and traditional RDBMS systems (which are deep in functionality)".

\begin{itemize}
  \item mongoDB implements a document-oriented structure, which is similar to
  KVP;
  \item mongoDB is written in C++, and therefore, can is available in any major
  platform, as well as offers a broad range of API drivers written in different
  languages such as Java, Python, Perl and Ruby; 
  \item mongoDB is open-source, with good community support and availability
  through mailing lists, freenode IRC channel, and commercial support through
  10gen company; 
  \item mongoDB has support to distributed systems properties such as
  Master-Slave replication, and features like Database Shards with
  auto-sharding based on shard keys.
\end{itemize}

The artifacts of the mongoDB are located in the thirdparty directory of the
NetBEAMS resources. However, the build script from the DSP Data Persistence
component can be executed to produce the persistence directory
structure for NetBEAMS, as described in figure
\ref{fig:dsp-persistence-system-dir}.

\begin{figure}[!h]
  \centering
  \includegraphics[scale=0.5]{../diagrams/dsp-persistence-system-dir}
  \caption{The osgi-intro Framework Layers}
  \label{fig:dsp-persistence-system-dir}
\end{figure}

Installation of mongoDB
- Explanation about mongoDB processes: mongod, mongos, mongo
- Specification of the shards, shard key, etc

    * The database instance is called "netbeams";
    * The database "netbeams" may contain different collections, categorized by
    the Sensor Content Type, that is, depending on how the DSP Component was
    described;

\subsection{Document-Oriented Data Model}

An example of the data is as folows (using the JSON syntax). Listing
\ref{file:mongodb-ysi-data-format} follows the strategy described on Section X,
where the entity representation is denormalized, and each of the properties of
the sensor is repeated in each instance of the data. The fact and transaction
times are expressed in milliseconds. The data element contains the complete
structure of the originating sensor such as the IP address. The collection of
instances of this document represents the sensor data type.

\subsection{Starting the Database}

\subsection{Stopping the Database}

\subsection{Execution Logs}

\section{Experiment Execution}

The experiments conducted using mongoDB can be summarized by the different
data access methods provided by the technology. Each subsection describes these
methods.

\subsection{Data Access Using Database Shell}

The mongoDB client can be started by using the following command. Make sure you
have started the mongoDB server before executing the mongoDB client.

mongo netbeams | tee output\underline{ }number\underline{ }date.log

Here, the iterative mongo client shell offers users to verify and navigate on a
given database and its collections. This first section shows the connection of
the mongo client to the database netbeams. It also highlights the query for the
collections available. During the experiment, the SondeDataContainer?
collection was created as related to the type from the DSP Messages for the YSI
Sonde.

The shell references to the mongoDB system can be found at
http://www.mongodb.org/display/DOCS/dbshell+Reference

\lstset{label=cmd:mongo,caption=Execution of mongo client}
\begin{lstlisting}
MongoDB shell version: 1.1.0-
url: netbeams
connecting to: netbeams
type "help" for help
> show collections
SondeDataContainer
system.indexes
>
> db.SondeDataContainer.count()
1000000
\end{lstlisting}

Then, the first verification of the data integrity is regarding the number of
elements created. Here, the first count() function on the collection returned
1000000.

An example about retrieving the first element of the collection can be done
using the findOne() function. It will return an element instance on the
JSONnotation.

\lstset{label=cmd:mongo-findone,caption=Querying the database: one item}
\begin{lstlisting}
> db.SondeDataContainer.findOne()
{
    "_id" : ObjectId("5d6f40078ec8074bba980900"),
    "message_id" : "020d82e1-18b2-4fe2-800c-a0f68e22ea86",
    "sensor" : {
        "ip_address" : "192.168.0.91",
        "location" : {
            "latitude" : 37.89155,
            "longitude" : -122.4464
        }
    },
    "time" : {
        "valid" : "Wed Nov 04 2009 02:30:12 GMT-0800 (PST)",
        "transaction" : "Sat Nov 21 2009 03:01:33 GMT-0800 (PST)"
    },
    "observation" : {
        "WaterTemperature" : 88.58,
        "SpecificConductivity" : 180.5,
        "Conductivity" : 167.3,
        "Resistivity" : 491.89,
        "Salinity" : 0.06,
        "Pressure" : 1.109,
        "Depth" : 2.642,
        "pH" : 7.11,
        "pHmV" : -42.9,
        "Turbidity" : 0.2,
        "ODOSaturation" : 72.4,
        "ODO" : 10.85,
        "Battery" : 2.7
    }
}
\end{lstlisting}

The query based on attributes can be done using the "dot" notation, as you
navigate through the JSON documents. Additionally, you can use the functions as
aggregated on the result of others. The example in listing
\ref{cmd:mongo-find-list} counts the number of documents with the key
"data.ph" equals to "5.64".

\lstset{label=cmd:mongo-find-list,caption=Execution of mongo client}
\begin{lstlisting}
> db.SondeDataContainer.find({"data.ph":5.64)}).count()
1226
\end{lstlisting}

The following example is the output of the first 2 documents from the same
previous query using the limit() function, as shown in figure
\ref{cmd:mongo-find-limit}.

\lstset{label=cmd:mongo-find-limit,caption=Query Element with specific
projection limiting the result set size}
\begin{lstlisting}
> db.SondeDataContainer.find({"data.ph":5.64}).limit(2)
{"_id" :  ObjectId( "d36f4007b7e7ac4a03c60000")  , "sensor_ip_address" : "192.168.0.136" , "message_id" : "7b6624d6-0ca1-4cba-a343-f166e88da73b"
, "transaction_time" : 1252845473412 , "fact_time" : 1252845346000 , "data" : {"temperature" : "45.01" , "sp_condition" : "37.6" ,
"condition" : "145.8" , "resistence" : "159.77" , "salinitude" : "0.0" , "pressure" : "0.391" , "depth" : "0.46" , "ph" : "5.64" ,
"pH_mv" : "-62.1" , "odo_sat" : "89.7" , "odo_condition" : "59.34" , "turbidity" : "0.0" , "battery" : "9.4"}}
{"_id" :  ObjectId( "d36f4007b7e7ac4a1fc80000")  , "sensor_ip_address" : "192.168.0.136" , "message_id" : "7b6624d6-0ca1-4cba-a343-f166e88da73b" ,
"transaction_time" : 1252845473412 , "fact_time" : 1252845346000 , "data" : {"temperature" : "46.71" , "sp_condition" : "60.8" ,
"condition" : "160.6" , "resistence" : "1399.4" , "salinitude" : "0.01" , "pressure" : "1.057" , "depth" : "2.485" , "ph" : "5.64" ,
"pH_mv" : "-16.3" , "odo_sat" : "58.8" , "odo_condition" : "19.29" , "turbidity" : "0.2" , "battery" : "9.2"}}
>
\end{lstlisting}

Other logs are located at
http://code.google.com/p/netbeams/downloads/list

\subsection{Data Access Using APIs}

The mongoDB server offers different drivers to access the data, as well as the
Web Services.

\begin{itemize}
  \item The Java tutorial on the drivers is located at
    http://www.mongodb.org/display/DOCS/Java+Tutorial. The driver is located on
    the NETBEAMS/versions/v2/thirdparty/mongodb directory;  
  \item The REST API can be called from any HTTP client. At the time of the
  editing, this feature is still under alpha version. Check the documentation
  athttp://www.mongodb.org/display/DOCS/Http+Interface for details 
\end{itemize}

The following HTTP GET Request method returns the first 5 documents in the collection:
http://127.0.0.1:28017/netbeams/SondeDataContainer/?limit=-5, as shown in
listing \ref{cmd:mongo-rest-request}.

\lstset{label=cmd:mongo-rest-request,caption=REST HTTP GET Request Example}
\begin{lstlisting}
HTTP/1.0 200 OK
x-action:
x-ns: netbeams.SondeDataContainer
Content-Type: text/plain;charset=utf-8

{
  "offset" : 0,
  "rows": [
    { "_id" : "156f4007e4c3b74a36ed3100", "sensor_ip_address" : "192.168.0.117", "message_id" : "08b02c08-9290-4517-9a28-c6ee7e16509a",
"transaction_time" : 1253557219486, "fact_time" : 1253557217000, "data" : { "temperature" : "31.44", "sp_condition" : "99.8", "condition" : "53.5",
"resistence" : "1157.08", "salinity" : "0.0", "pressure" : "1.066", "depth" : "0.161", "ph" : "1.08", "pH_mv" : "-82.0", "odo_sat" : "40.3",
"odo_condition" : "56.85", "turbidity" : "0.2", "battery" : "8.2" } } ,
    { "_id" : "156f4007e5c3b74a37ed3100", "sensor_ip_address" : "192.168.0.117", "message_id" : "08b02c08-9290-4517-9a28-c6ee7e16509a",
"transaction_time" : 1253557219486, "fact_time" : 1253557217000, "data" : { "temperature" : "37.83", "sp_condition" : "176.3", "condition" : "2.6",
"resistence" : "1324.97", "salinity" : "0.01", "pressure" : "1.36", "depth" : "1.564", "ph" : "0.12", "pH_mv" : "-23.5", "odo_sat" : "104.5",
"odo_condition" : "19.44", "turbidity" : "0.1", "battery" : "5.0" } } ,
    { "_id" : "156f4007e5c3b74a38ed3100", "sensor_ip_address" : "192.168.0.117", "message_id" : "08b02c08-9290-4517-9a28-c6ee7e16509a",
"transaction_time" : 1253557219486, "fact_time" : 1253557217000, "data" : { "temperature" : "74.3", "sp_condition" : "84.0", "condition" : "104.7",
"resistence" : "4089.13", "salinity" : "0.01", "pressure" : "1.222", "depth" : "2.788", "ph" : "6.56", "pH_mv" : "-78.1", "odo_sat" : "40.0",
"odo_condition" : "6.02", "turbidity" : "0.3", "battery" : "3.2" } },
  "total_rows" : 3 ,
  "query" : {} ,
  "millis" : 0
}
\end{lstlisting}

Data visualisation tools for mongoDB is slowly being developed by open-source
developers. The next picture shows the database "netbeams" and the collection
"SondeDataContainer  ?" being rendered by futon4mongodb, one of the
open-source tools developed to visualise mongoDB data.

\subsection{Visualizing Data on The Browser}

Futon4Mongo is an open-source software adapted from CouchDB \ref{couchdb}.
First, the collections list is shown. Figure
\ref{fig:view-collections-instance-browser-futondb} shows one single collection
for the YSI data containing 1 million objects.

\begin{figure}[h]
  \centering
  \includegraphics[scale=0.3]{../diagrams/view-collections-instance-browser-futondb}
  \caption{Viewing a partial list of data using the Futon for CouchDB/MongoDB}
  \label{fig:view-collections-instance-browser-futondb}
\end{figure}

The collection is composed by instances of documents of representing the YSI
Sonde type, being indexed by a document ID and the values being the keys
defined in the previous chapter, as shown in figure
\ref{fig:view-collected-data-list-browser-futondb}.

\begin{figure}[h]
  \centering
  \includegraphics[scale=0.3]{../diagrams/view-collected-data-list-browser-futondb}
  \caption{Viewing a partial list of data using the Futon for CouchDB/MongoDB}
  \label{fig:view-collected-data-list-browser-futondb}
\end{figure}

By clicking in one of the items, one can view the instance of the document, as
it is shown in figure \ref{fig:view-collections-instance-browser-futondb}.

\begin{figure}[h]
  \centering
  \includegraphics[scale=0.3]{../diagrams/view-collected-data-instance-browser-futondb}
  \caption{Viewing an instance of collected data using the Futon for
  CouchDB/MongoDB}
  \label{fig:view-collected-data-instance-browser-futondb}
\end{figure}

\subsection{Exporting Data to Spreadsheets}

mongoDB has an export facility shell called mongoexport. It can export the data
in JSON format or CSV. One may also write its own export tool in any of the
languages such as Java, PHP, Python, Perl, Ruby, among others. A list of the
existing drivers in different languages is provided
athttp://www.mongodb.org/display/DOCS/Drivers. The following command can be
executed to have the exported version of the data in CSV (read the help output
of the command for details).

An example to export the data in CSV format can be seen in listing
\ref{cmd:mongoexport}, and an example containing 1 million objects can be
downloaded at
http://netbeams.googlecode.com/files/experiment-1000000-data-exported-20090913-053538.csv.tar.gz.

\lstset{label=cmd:mongoexport,caption=Command to export data in CSV format}
\begin{lstlisting}
mongoexport -d netbeams -c SondeDataContainer --dbpath ./data/ --csv -f "_id,sensor_ip_address,transaction_time,fact_time,
data.temperature,data.sp_condition,data.condition,data.resistence,data.salinitude,data.pressure,data.depth,data.ph,data.pH_mv,data.odo_sat,
data.odo_condition,data.turbidity,data.battery" -o sonde-data-exported.csv
\end{lstlisting}

\subsection{Exporting Data to OPeNDAP Format}

By using any programming language driver, one can export data in the OPeNDAP
format.

@@

\lstinputlisting[language=Java,label=file:random-ysi-data-generator,caption=Experiment
Implementation]{../../../../netbeams/versions/v2/apps/osgi-bundles/dsp/DSPSondeProducer/src/org/netbeams/dsp/ysi/SondeTestData.java}

@@

\lstset{label=file:mongodb-ysi-data-format,caption=JSON representation of the
data produced by a YSI sensor on mongoDB}
\begin{lstlisting}
{
    "_id" : ObjectId("5d6f40078ec8074bba980900"),
    "message_id" : "020d82e1-18b2-4fe2-800c-a0f68e22ea86",
    "sensor" : {
        "ip_address" : "192.168.0.91",
        "location" : {
            "latitude" : 37.89155,
            "longitude" : -122.4464
        }
    },
    "time" : {
        "valid" : "Wed Nov 04 2009 02:30:12 GMT-0800 (PST)",
        "transaction" : "Sat Nov 21 2009 03:01:33 GMT-0800 (PST)"
    },
    "observation" : {
        "WaterTemperature" : 88.58,
        "SpecificConductivity" : 180.5,
        "Conductivity" : 167.3,
        "Resistivity" : 491.89,
        "Salinity" : 0.06,
        "Pressure" : 1.109,
        "Depth" : 2.642,
        "pH" : 7.11,
        "pHmV" : -42.9,
        "Turbidity" : 0.2,
        "ODOSaturation" : 72.4,
        "ODO" : 10.85,
        "Battery" : 2.7
    }
}
\end{lstlisting}

### 2009-11-27 docs/marcello-dissertation/thesis/chapter3-taxonomic-classification-of-persistence-storage-for-sensor-networks.tex
Major revision of the thesis structure, and review of chapters, moving chapter 7 to the appendix, and adding the first revision of the abstract in the preamble.
Chap 3. Reviewed the structure, misspelled words, sentences, etc;

Chap 4. Changes to the structure of the case study. The documentation regarding netbeams was pushed to the last chapter: appendix;

Chap 5. The selection of the technology follows the review with Prof. Puder. One section describes each of the technologies briefly, and details their characteristics during the analysis of each of the taxonomies;

Chap 6. Describes the imp

Different mechanisms are used for persisting data collected from sensor
networks, as described by the section regarding the state of the art in sensor
networks and different approaches for persistence. This section analyzes these
approaches and presents the different approaches through 
taxonomy\footnote{Taxonomy is the practice and science of classification, by
using taxonomic units known as taxa (singular taxon)}, which will be used
throughout this work to evaluate different existing database systems
implementations, taking into account different requirements and specifications
of different sensor networks.

@@

No matter the type of resource management used, the first taxonomy presented is
related to the purpose of the use of the collected sensor data. That is, when
and how the collected data is going to be used as described in section
\ref{sec:sn-data-purpose}.

@@

Directly related to the purpose of data, the mechanism for data storage
described in section \ref{sec:sn-infrastructure} depends on the purpose of
the use of the collected data. The related taxon is depicted by figure
\ref{fig:taxonomy-data-location}, and described as follows:

@@

\begin{itemize}
  \item \textbf{Local Storage}: The data can be temporarily persisted
   in-memory, or a secondary memory device in a local storage device. It is
  directly related to the real-time data stream taxon, since it occurs at
  the place of the sensor device;
  \item \textbf{External Storage}: The data is usually located in an
  external storage device with a data management system such as a database
  system.
  \item \textbf{Data-Centric Storage}: the collected data can arranged in a
  categorized fashion as described by \cite{sn-storage03}. This taxon of
  system allows the data to be partitioned by a specific property. Nowadays,
  the strategy is often related to database partition \cite{db-partition}
  and database sharding approaches \cite{db-shard01} \cite{db-shard02}.
\end{itemize}

@@

One of the major challenges in the development of a persistence storage for
sensor network is related to the data model chosen to represent the collected
data. The addition of new sensor devices may trigger refactoring of the
existing model or not, depending on how the data model accommodates the data.
Different data models are reported to be used in the literature, as described
in section \ref{sec:data-models}. For this reason, taxonomy related to the
data model depicts two different classes of data models, as shown in figure
\ref{fig:taxonomy-data-model}.

@@

\begin{itemize}
  \item \textbf{Schema-Dependent Models}: this type of data model require
  the definition of a master schema that describes the data through a regorous
  data modeling process prior to the use of the data. In addition, an addition
 to a new sensor device, and consequently a new data entity, requires a
 re-evaluation of the existing data schema. For instance, relational model is
 governed by the relational algebra model, while the structured models such as
 XML needs a definition of an XML Schema.
  \item \textbf{Schema-less Models}: contrary to the previous taxon, 
  schema-less data models does not require the definition of a schema or master
  table definition and relationships. Instances of such data model is the KVP
  and the document-oriented models, whose usage in wireless sensor networks
  have not been reported.
\end{itemize}

@@

No matter the type of data model used, the support for Data Provenance as shown
in section \ref{sec:sn-provenance} is as important as the data being collected.
For this reason, the data model taxonomy must have support for the data types
that support the use of provenance-related attributes in the collected sensor
data. In this way, Figure \ref{fig:taxonomy-data-provenance} describes the data types
for the metadata support.

@@

\begin{itemize}
  \item \textbf{What: Data Identity}: the data type that uniquely identifies
  the collected sensor dat;
  \item \textbf{When: Time Dimension}: the data type that identifies the
  collected data;
  \item \textbf{Where: Data Location}: an optional data type that identifies
  where the data came from.
\end{itemize}

@@

The query processing mechanism in sensor networks is directly related to how
and where the collected data is used and stored, as shown in the previous
section. As a consequence, the following taxonomy taxon is related to the
different types of query processing mechanisms described in section
\ref{sec:query-process}, and can be seen in figure
\ref{fig:taxonomy-query-mechanism}.

@@

\begin{itemize}
  \item \textbf{In-Network Query Processing}: the sensor node receives a
  request to the properties of the sensor, and returns the current values
  related to them \cite{sn-intro01}.
  \item \textbf{Centralized Query Processing}: as it is done in regular single
  server relational databases, the data is collected to a centralized data
  sink in order to be used.
\end{itemize}

@@

Different approaches of data storage can also be related to the performance in
the database system used, and how its architecture provides the data storage
solution. For example, one important problem related to the centralized
network sink in sensor networks is the presence of an unavoidable creation of a
point of traffic concentration at the data collector or network nodes as
described by \cite{sn-storage02}. In this way, in order to mitigate the
problems associated with the so-called ``Funneling Effect", figure
\ref{fig:taxonomy-database-architecture} summarizes the different capabilities
of different types of database systems.

@@

\begin{itemize}
  \item \textbf{Centralized System}: this is a database system that runs in
  one single centralized host, being the main point of data traffic as read
  and writes \cite{sn-intro01}.
  \item \textbf{Distributed System}: a database system that can be composed by
  more than one node. Usually, the use of the paradigm of Master-Slave
  mechanisms can be used by implementing ``Replication". In addition to it,
  ``Database Sharding" is used to portion the data related to a given shard
  key.
\end{itemize}

>>>>

Different mechanisms are used for persisting data collected from sensor
networks, as described by the section regarding the state of the art in sensor
networks and different approaches for persistence. This section analyzes these
approaches and presents the different approaches through
taxonomy\footnote{Taxonomy is the practice and science of classification, by
using taxonomic units known as taxa (singular taxon)}, which were used
throughout this work to evaluate different existing database systems
implementations, taking into account different requirements and specifications
of different sensor networks.

@@

No matter the resource management used, the first taxonomy presented relates to
the purpose of the use of the collected sensor data. That is, when and how the
collected data is going to be used as described in section
\ref{sec:sn-data-purpose}.

@@

Directly related to the purpose of data, the procedure for data storage
described in section \ref{sec:sn-infrastructure} depends on the purpose of the
use of the collected data. Figure \ref{fig:taxonomy-data-location}, shows the
related taxon and can be described as follows:

@@

\begin{itemize}
  \item \textbf{Local Storage}: The data can be persisted in a temporarily
  memory address, or a secondary memory device in a local storage device. It is
  directly related to the real-time data stream taxon, since it occurs at the
  place of the sensor device;
  \item \textbf{External Storage}: The data is usually located in an external
  storage devicewith a data management system such as a database system;
  \item \textbf{Data-Centric Storage}:the collected data can organized in a
  categorizedmanner as described by \cite{sn-storage03}. This taxon regarding
  the system allows the data to bepartitioned by a particular property.
  Nowadays, this strategy is often related to database partition
  \cite{db-partition}, and database sharding approaches \cite{db-shard-intro,
  db-shard-discussion}.
\end{itemize}

@@

One of the fundamental challenges in the development of a persistence storage
in sensor network is related to the data model chosen to keep the collected
data. The introduction of new sensor devices may cause refactoring of the
existing model or not, depending on how the data model accommodates the data.
Different data models are reported to be used in the literature, as described
in section \ref{sec:data-models}. For this reason, taxonomy related to the
data model depicts two different classes of data models, as shown in Figure
\ref{fig:taxonomy-data-model}.

@@

\begin{itemize}
  \item \textbf{Schema-Dependent Models}: this data model requires the
  definitionof a master schema that describes the data through a rigorous data
  modeling process prior to the use of the data. In addition, an addition to a
  newsensor device, and hence a new data entity, requires a re-evaluation ofthe
  existing data schema. For example, relational model is governed by
  therelational algebra model \cite{relational-model}, while the structured
  models such as XML \cite{xml} needs adefinition of an XML Schema
  \cite{xml-schema};
  \item \textbf{Schema-less Models}: contrary to the former taxon, schema-less
  data models do not require the definition of a data schema or a master table
  definition, along with its table relationships, if necessary. Instances of
  such data model is the Key-Value Pair \cite{db-kvp} and its variation
  document-oriented models \cite{db-is-rdbs-dommed}, whose use in wireless
  sensor networks has not been reported.
\end{itemize}

@@

No matter the data model used, the support for Data Provenance as shown in
section \ref{sec:sn-provenance} is as valuable as the data being collected.
For this reason, the data model taxonomy must have support for the data types
that support the use of provenance-related attributes in the collected sensor
data. In this way, Figure \ref{fig:taxonomy-data-provenance} describes the
data types for the metadata support.

@@

\begin{itemize}
  \item \textbf{What: Data Identity}: the data type that uniquely identifies
  the collected sensor data;
  \item \textbf{When: Time Dimension}: the data type that identifies the
  collected data;
  \item \textbf{Where: Data Location}: an optional data type that identifies
  from where the data originated.
\end{itemize}

@@

The query processing mechanism in sensor networks is directly related to how
and where the collected data is used and stored, as shown in the previous
section. As a consequence, the following taxonomy taxon is related to the
different types of query processing mechanisms described in section
\ref{sec:query-process}, and can be seen in Figure
\ref{fig:taxonomy-query-mechanism}.

@@

\begin{itemize}
  \item \textbf{In-Network Query Processing}: the sensor node receives a
  request to theproperties of the sensor, and returns the current values
  related to them \cite{sn-intro01};
  \item \textbf{Centralized Query Processing}: as developed in single server
  relational databases, the data is collected to a centralized data sink in
  order to be used.
\end{itemize}

@@

Different approaches of data storage can also be related to the performance in
the database system used, and how its architecture provides the data storage
solution. For example, one crucial problem related to the centralized network
sink in sensor networks is the presence of an ineluctable develoment of a point
of traffic concentration at this node \cite{sn-storage02}. In this way, in
order to reduce the problems associated with the so-called ``Funneling Effect",
Figure \ref{fig:taxonomy-database-architecture} summarizes the different
capabilities of different types of database systems.

@@

\begin{itemize}
  \item \textbf{Centralized System}: a database system that runs in one single
  centralized host, being the focal purpose of data traffic as read and writes 
  \cite{sn-intro01};
  \item \textbf{Distributed System}: a database system that can be composed by
  more thanone node. Usually, the use of the pattern of Master-Slave
  mechanisms can beused by implementing ``Replication". In addition to it,
  ``Database Sharding" is used to portion the data related to a given shard key.
\end{itemize}
